Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending

### 2017-09-12 paper.Rmd
better sample selection, now getting full citations data

---
title: Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending
author:
  - name: Joe Brew
    email: joebrew@economicsofmalaria.com
    affiliation: isglobal,cism
    footnote: Corresponding Author
  - name: Menno Pradhan
    email: m.p.pradhan@vu.nl
    affiliation: vu,uva
  - name: Elisa Sicuri
    email: elisa.sicuri@isglobal.org 
    affiliation: isglobal,icl,cism
address:
  - code: isglobal
    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
  - code: icl
    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
  - code: cism
    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
  - code: vu
    address: Faculty of Economics and Business Administration, VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
  - code: uva
    address: Faculty of Economics and Business, Section Microeconomics, University of Amsterdam, REC E, Roetersstraat 11, Amsterdam, Netherlands
abstract: |
  Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to make fully informed decisions. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a systematic survey of malaria researchers, we query perceptions regarding the likelihood and time frame of eradication, as well as the perceived chief areas needed for improvement in order for eradication to be achieved. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise. Our results, which suggest a disconnect between public institutional discourse and private opinion, serve as a barometer of professional perspectives, and serve to highlight areas where attention is most needed for eradication to be achieved.

bibliography: library.bib
output: rticles::elsevier_article
---

@@

# Packages (load child-specific packages in their respective files)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(RColorBrewer)
# library(Hmisc)
# library(Rmisc)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.align = 'center')
```

```{r}
# Get publication theme for charts
source('theme_publication.R')

# Send emails
source('code/recruit.R', chdir = TRUE)

# Load data
file_name <- paste0('snapshot_', Sys.Date(), '.RData')
if(!file_name %in% dir('snapshots')){
  # Fetch data
  library(gsheet)
  url_responses<- 'https://docs.google.com/spreadsheets/d/1yeQ4TR73mdnDABPZ-ZEg3DAFu2Bko5NJ_nkMcNKr2lo/edit?usp=sharing'
  df <- gsheet2tbl(url_responses)

  # Manually rename variables
  names(df) <- tolower(gsub('.', '_', names(df), fixed = TRUE))
  names(df)[4:17] <-
    c('area',
      'years_10',
      'years_20',
      'years_30',
      'years_40',
      'years_50',
      'years',
      'governance',
      'health_workforce',
      'health_financing',
      'health_technologies',
      'health_information',
      'service_delivery',
      'comments')

  # Save a snapshot
  save(df, 
       file = paste0('snapshots/',
                     file_name))
} else {
  load(paste0('snapshots/',
              file_name))
}
```

```{r}
# Clean up survey responses
# Make NA all empty responses
df$comments[df$comments == ''] <- NA
df$email_address[df$email_address == ''] <- NA
# Flag duplicates
df$duplicate <- duplicated(df$email_address)
# Remove duplicates
df <- df[is.na(df$email_address) |
           !df$duplicate,]
# Create area dummies
df$area <- tolower(df$area)
df$area <- gsub('/', ', ', df$area)
df$area <- gsub('(', '', df$area, fixed = TRUE)
df$area <- gsub(')', '', df$area, fixed = TRUE)
# Perform some replacements
df$area[df$area == 'biochemistry'] <- 'chemistry'
df$area[df$area == 'medical entomology'] <- 'entomology'
df$area[df$area == 'vector biology'] <- 'biology'

areas <- tolower(paste0(sort(unique(df$area)), collapse = ', '))
areas <- sort(unique(unlist(strsplit(x = areas, split = ','))))
areas <- trimws(areas, which = 'both')
old_areas <- areas

# Remove those areas for which there are fewer than 5 of a kind
areas_df <- data.frame(area = areas)
areas_df$count <- NA
for (i in 1:nrow(areas_df)){
  areas_df$count[i] <-
    length(which(grepl(areas_df$area[i], 
          df$area)))
}
areas_df <- areas_df %>%
  filter(area != '',
         count >= 5)
areas <- areas[areas %in% areas_df$area]

for (j in 1:length(areas)){
  df[,paste0(
    'area_',
    areas[j]
  )] <- NA
}
for (j in 1:length(areas)){
  df[,paste0(
    'area_',
    areas[j]
  )] <- grepl(areas[j], df$area)
}

# Get number of areas
# Average number of areas
area_counter <- function(x){
  result <- unlist(gregexpr(',', x))
  if(result[1] == -1){
    return(1)
  } else {
    return(length(result) + 1)
  }
}
df$n_areas <- sapply(df$area, area_counter)

# Make years numeric
df$years <- tolower(df$years)
# more needs to be done here

# Make comments standardized
df$comments <- tolower(df$comments)

# Clean up names
names(df) <- tolower(gsub('.', '_', names(df), fixed = TRUE))
names(df) <- gsub(' ', '_', names(df))
names(df) <- gsub('-', '_', names(df))
names(df) <- gsub('&', 'and', names(df), fixed = TRUE)
names(df) <- gsub('__', '_', names(df))

df <- data.frame(df)

# Clean up missing responses
for (i in seq(10, 50, 10)){
  df[,paste0('years_', i)][df[,paste0('years_', i)] == ''] <- NA
  df[,paste0('years_', i)] <-
    factor(df[,paste0('years_', i)],
           levels = c('0%',
                      '1-20%',
                      '21-40%',
                      '41-60%',
                      '61-80%',
                      '81-99%',
                      '100%'))
}

# Clean up years
df$years[df$years %in% c('', '10%',
                         'cannot say',
                         'depends on approach and commitment',
                         "human species may be eradicated but zoonotic species may come up",
                         "i don't know - too many unknown variables",
                         "it depends...hard to predict.",
                         "i think we have to be humble. malaria eradication was promised a long time ago...",
                         "it is depend on global stability and climate change",
                         "l don't know but it will remain big health problem",
                         "not sure - depends on when the world begins to take the right steps towards eradication.",
                         'not year but want to')] <- NA
df$years[grepl('never|nevet|ever', df$years)] <- 50
df$years <- gsub(' years| yrs|years|yrs| yeras|about ', '', df$years)
df$years[df$years %in% c('100+',
                         '≥100 to a global zero',
                         '> 50',
                         '50-100',
                         '50 ish',
                         '50yrs',
                         '60 (one more generation)',
                         '75a',
                         'at least 400',
                         "cannot be fully erdicated-some local eradication can be achieved within 50 ",
                         'it will not be eradicated',
                         'more than 50',
                         'unlikely',
                         'i doubt it can be eradicated at all')] <- 50
df$years[df$years == '10-15'] <- 12.5
df$years[df$years == '20-30'] <- 25
df$years[df$years == '40-50'] <- 45
df$years <- as.numeric(as.character(df$years))
df$years[df$years > 50] <- 50

# Clean up the rankings
bb <- c('governance', 'health_workforce',
        'health_financing', 'health_technologies',
        'health_information', 'service_delivery')
for (i in 1:nrow(df)){
  sub_df <- df[i,bb]
  new_values <- sub_df / sum(sub_df, na.rm = TRUE) * 21
  new_values[new_values > 6] <- 6
  df[i,bb] <- new_values
}

```{r}
# Results cleanup
```

@@

>>>>

---
title: Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending
author:
  - name: Joe Brew
    email: joebrew@economicsofmalaria.com
    affiliation: isglobal,cism
    footnote: Corresponding Author
  - name: Menno Pradhan
    email: m.p.pradhan@vu.nl
    affiliation: vu,uva
  - name: Jacqueline Broerse
    email: j.e.w.broerse@vu.nl
    affiliation: vu
  - name: Elisa Sicuri
    email: elisa.sicuri@isglobal.org 
    affiliation: isglobal,icl,cism
address:
  - code: isglobal
    address: Barcelona Institute for Global Health, c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain
  - code: icl
    address: Imperial College London, South Kensington Campus, London SW7 2AZ, U.K.
  - code: cism
    address: Centro de Investigação em Saúde de Manhiça, Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique
  - code: vu
    address: Faculty of Economics and Business Administration, VU University Amsterdam, De Boelelaan 1105, 1081 HV Amsterdam, Netherlands
  - code: uva
    address: Faculty of Economics and Business, Section Microeconomics, University of Amsterdam, REC E, Roetersstraat 11, Amsterdam, Netherlands
abstract: |
  Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to make fully informed decisions. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a systematic survey of malaria researchers, we query perceptions regarding the likelihood and time frame of eradication, as well as the perceived chief areas needed for improvement in order for eradication to be achieved. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise. Our results, which suggest a disconnect between public institutional discourse and private opinion, serve as a barometer of professional perspectives, and serve to highlight areas where attention is most needed for eradication to be achieved.

bibliography: library.bib
params:
  get_new_data: False
output: rticles::elsevier_article
---

@@

# Packages (load child-specific packages in their respective files)
library(knitr)
library(xtable)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.align = 'center')
options(xtable.comment = FALSE)

# Get data
source('prepare_data.R')
```

@@

## Sample selection and de-biasing of results

This section is unfinished. Below are regression outputs for both OLS and Heckman's 2 step model. I will update this soon.

```{r, results = 'asis'}
stargazer(ols1, heck1, ml1,
          header = FALSE,
          # se=list(cse(ols1),NULL,NULL),
          title="Regression on years until eradication", 
          type="latex",
          df=FALSE, digits=4)
# summary(ols1)
# summary(heck1)
